\textbf{ Matching the forecast horizon with the relevant ecological processes }

### 2019-09-13 doc/manuscript.tex
Edited Ethan's spatiotemporal addition

Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...\comment{Ethan, what do you have in mind here? Probably something more than just working with spatiotemporal datasets and accounting for autocorrelation. Mevin has lots of papers with some sort of mechanistic modeling of spatiotemoral processes like dispersal and colonization. Should we say that the datasets needed to train this kind of model are rare, but that might be changing with more and more remote sensing data that is extensive in space and time? }

>>>>

Another potential approach for intermediate scale forecast horizons is to use models that incorporate both spatial and temporal information simultaneously. Spatiotemporal modeling approaches are being developed to study patterns and processes of interest to ecological forecasters, such the spread of an invasive species or population status of a threatened species  \citep{Wikle2003,Williams2017,Schliep2018}. Because these models include both fast processes, such as births and deaths, and slower processes, such as colonization and extinction dynamics, they have the potential to make better predictions at intermediate forecast horizons than purely spatial or temporal models. However, these spatiotemporal models have rarely been used in a forecasting context, due to a combination of data limitation and computational challenges. Many data sources contain either spatial or temporal variation, but not both, and when spatiotemporal datasets are available they often involve irregular sampling, creating challenges for modeling. Fitting and generating predictions from spatiotemporal models is also computationally intensive, especially with large datasets \citep{McDermott2017}. Fortunately, thanks to large-scale monitoring efforts from remote sensing platforms, the National Ecological Observatory Network (https://www.neonscience.org/), and community science projects (e.g., eBird), large scale spatiotemporal data is increasingly available. In addition, new methods for spatiotemporal forecasting are being developed that address existing computational challenges \citep{McDermott2017}, and access to high performance computing resources is increasingly common. Given these developments, future ecological forecasting efforts should explore spatiotemporal approaches and assess whether they improve predictions at intermediate time scales relative to traditional spatial or temporal models.
